---
title: "janna's portfolio"
author: "janna"
date: "2023-02-15"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true 

---


```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = FALSE, message = FALSE, warning = FALSE, results = FALSE)
library(ggplot2)
library(tidyverse)
library(spotifyr)
library(plotly)
library(flexdashboard)
```

### INTRODUCTION 

"When you are young they assume you know nothing" 


*** 
 In my musical corpus, I seek to answer the question of what is "sad girl music", what does it mean to be a "sad girl" singer, songwriter and producer? In aiming to answer such questions, I've amassed a musical corpus which reflects the most popular trends and artists within the sad girl genre, such as Taylor Swift's pandemic sister albums "folklore" and "evermore", as well as unique songwriters which have most recently gained popularity, such as Phoebe Bridgers with her "Stranger in the alps" and "Punisher" albums. Additionally, I've added well-known pillars of the genre, such as Lana del Rey and Mitski, as well as artists appreciated for their longevity and whose older songs have remained iconic - most specifically, Fiona Apple. Additionally, I aim to uncover what it means for an artist known for her "sad girl music" to diverge from said patterns and apperently go through her "healing era" - this can be observed in the commercial trends of artists such as Lorde and fka twigs, both included in the corpus. Finally, newcomers and new sounds and trends are also included in the corpus, through Lizzy McAlpine, Gracie Abrams and Ethel Cain, whose comparisons will help us understand the new and up and coming "sad girl" music, and how much it reflects the core ideas already installed by more popular artists.


### What Is Sad Girl Music, and What Is Not? 


```{r corpus, echo=FALSE}

sadgirl <- get_playlist_audio_features(" ", "1aCB5tYDsZoHJLH6wXNd7T")
notsadgirl <- get_playlist_audio_features (" ", "1dj5yWeujg3pPiIMEDakmm")
corpus <-
  bind_rows(
    sadgirl |> mutate(category = "Sad Girl Music"),
    notsadgirl |> mutate(category = "Not Sad Girl Music")
  )





corpus |>                    # Start with awards.
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) |>
  ggplot(                     # Set up the plot.
    aes(
      x = valence,
      y = energy,
      size = loudness,
      colour = mode,
      label = track.name
    )
  )+
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(                  # Add text labels from above.
    aes(
      x = valence,
      y = energy,
      label = label,
    ),
    data = 
      tibble(
        label = c(" ", " "),
        category = c("Not Sad Girl Music", "Sad Girl Music"),
        valence = c(0.090, 0.123),
        energy = c(0.101, 0.967),
      ),
    colour = "black",         # Override colour (not mode here).
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "bottom",         # Align bottom of label with the point.
    nudge_x = -0.05,          # Nudge the label slightly left.
    nudge_y = 0.02            # Nudge the label slightly up.
  ) +
  facet_wrap(~ category) +    # Separate charts per playlist.
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),   # Use grid-lines for quadrants only.
    minor_breaks = NULL       # Remove 'minor' grid-lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),
    minor_breaks = NULL
  ) +
  scale_colour_brewer(        # Use the Color Brewer to choose a palette.
    type = "qual",            # Qualitative set.
    palette = "Paired"        # Name of the palette is 'Paired'.
  ) +
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud.
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Valence",
    y = "Energy",
    colour = "Mode",
    label = "Track name"
  )
ggsave("corpus.pdf", plot=last_plot(), width=15, height = 10)
ggplotly(plot=last_plot())

```

